Translation Elicitation

### 2021-09-29 Rmd/report.Rmd
Remove big files

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE, 
  cache.extra = knitr::rand_seed,
  out.width = "100%",
  results = "asis",
  dev.args = list(png = list(type = "cairo"))
)
options(knitr.kable.NA = '-',
        ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.continuous.fill = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.continuous.colour = wesanderson::wes_palettes$Darjeeling1
)

@@

# set params
source(here("R", "utils.R"))

# import data
stimuli <- readRDS(here("Data", "stimuli.rds"))
participants <- readRDS(here("Data", "participants.rds"))
responses <- readRDS(here("Data", "responses.rds"))
jtrace <- readRDS(here("Data", "jtrace.rds"))

# import models
fit_jtrace <- readRDS(here("Results", "fit_jtrace.rds"))
fit_responses <- readRDS(here("Results", "fit_responses_4.rds"))
loo <- readRDS(here("Results", "loo.rds"))

@@

# TRACE simulations

## TRACE vs. human data {.tabset .tab-pill}

### TRACE activation vs. responses

```{r jtrace_all_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(jtrace, aes(x = activation, y = pct_participant)) +
  geom_smooth(colour = wes_palettes$Darjeeling1[1],
              fill = wes_palettes$Darjeeling1[1],
              alpha = 0.5) +
  geom_point(shape = 1, stroke = 1, alpha = 0.5) +
  labs(
    x = "jTRACE activation value for the word",
    y = "Participants who responded with the word"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_custom() +
  theme(
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  ) +
  ggsave(here("Figures", "trace_average.png"))

### By items

```{r jtrace_word_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# plot by word
jtrace %>% 
  mutate(input_language = paste0(input, " (", language, ")")) %>% 
  ggplot(aes(x = activation, y = pct_participant)) +
  facet_wrap(~input_language) +
  geom_smooth(colour = wes_palettes$Darjeeling1[1],
              fill = wes_palettes$Darjeeling1[1],
              alpha = 0.5) +
  geom_point(shape = 1, stroke = 1, alpha = 0.5) +
  labs(
    x = "jTRACE activation value for the word",
    y = "% participants who responded with the word",
    title = "jTRACE activation against % participants who gave word as response"
  ) +
  theme_custom() +
  theme(
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

## Regression model {.tabset .tab-pill}

### Fixed coefficients

```{r jtrace_fixed_table, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
summary(fit_jtrace$fit)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_number(2:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("pthn", "frequency", "lv", "pthn:lv") | str_detect(term, "activation"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  )
```

```{r jtrace_fixed_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%"}
fit_jtrace$fixed %>%
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4)) %>% 
  ggplot(aes(.value, fill = .variable)) +
  facet_wrap(~.variable, scales = "free") +
  geom_vline(xintercept = 0, colour = "black") +
  stat_halfeye() +
  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
  theme_custom() +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
    axis.text = element_text(size = 9)
  )
```

## Candidate pool

```{r jtrace_candidates, echo=FALSE, fig.height=30, message=FALSE, warning=FALSE, paged.print=FALSE}

stimuli_trace <- stimuli %>% 
  as_tibble() %>% 
  select(group, input = word1, target = word2, phon_trace)

# process TRACE data ----
trace <- map(
  list.files(here("Data", "jtrace"), full.names = TRUE),
  function(x){
    read_xlsx(x) %>% 
      pivot_longer(-cycle, names_to = "response", values_to = "activation")
  }
) %>%
  set_names(list.files(here("Data", "jtrace"))) %>% 
  bind_rows(.id = "target") %>% 
  separate(target, c("input", "language"), sep = "_") %>% 
  mutate(language = str_remove(language, ".xlsx"),
         language = str_to_sentence(language))

# get top candidates
trace_top <- trace %>% 
  group_by(input, language, response) %>% 
  summarise(
    activation = max(activation, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  group_by(input, language) %>% 
  slice_max(activation, n = 10) %>% 
  ungroup() %>% 
  mutate(group = ifelse(language=="Catalan", "ENG-CAT", "ENG-SPA")) %>% 
  select(-language) %>% 
  # add experimental trials
  left_join(stimuli_trace) %>% 
  mutate(correct = response==phon_trace) %>% 
  select(group, input, response, activation, correct)

# experimental results ----
responses_jtrace <- responses %>% 
  filter(valid_response, group %in% c("ENG-CAT", "ENG-SPA")) %>% 
  rename(response = input_text, input = word) %>% 
  count(group, input, response, correct) %>% 
  group_by(group, input) %>%
  mutate(total = sum(n), activation = n/total) %>% 
  ungroup() %>% 
  group_by(group, input) %>% 
  slice_max(order_by = activation, n = 10) %>% 
  ungroup() %>% 
  select(group, input, response, correct, activation) %>% 
  filter(input %in% trace_top$input)

# merge everything
d_jtrace <- list(TRACE = trace_top, Participants = responses_jtrace) %>% 
  bind_rows(.id = "source") %>% 
  mutate(
    # to reorder variable within panels
    input_label = paste0(input, " (", group, ")"),
    response_reordered = reorder_within(response, activation, input_label)
  )

ggplot(d_jtrace, aes(response_reordered, activation, fill = correct)) +
  facet_wrap(input_label~source, scales = "free_y", ncol = 3) +
  geom_col(colour = "white") +
  labs(x = "Top 10 candidates", y = "Max. TRACE activation") +
  coord_flip() +
  scale_x_reordered() +
  # scale_y_continuous(limits = c(0, 1)) +
  theme_bw() +
  theme(
    legend.position = "none"
  )

@@

>>>>

```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE, 
  cache.extra = knitr::rand_seed,
  out.width = "100%",
  results = "asis",
  dev.args = list(png = list(type = "cairo"))
)
options(knitr.kable.NA = '-',
        ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.continuous.fill = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.continuous.colour = wesanderson::wes_palettes$Darjeeling1
)

@@

# import data
tar_load_globals()
tar_load(stimuli)
tar_load(participants)
tar_load(responses)
tar_load(model_fits)
tar_load(model_loos)

theme_set(theme_github())

@@

@@

# {-}
